Contrastive Optimization -- Preliminaries

### 2025-01-03 paper/main.tex
reverted ablated bce formulation to only constraint target-class cams

>>>>

\subsubsection{Ablated Binary Cross-Entropy}

For sigmoid-activated binary / multilabel classification tasks, we leverage similar principles to defined an ablation to binary cross-entropy. Since we do not have a contrastive process in softmax-activation, this definitions relies only on HiResCAMs.

\begin{theorem}\label{abce}
	We can integrate backgroud invariance to the definition of binary cross-entropy using the following formulation:
	\begin{gather}
		H(\theta, \mathcal{D}) = - y_i \log \left(\sigma\left(h \odot \sum_{j,k}\tilde{\mathcal{\bm{A}}}_{i,j,k}^{\text{HiResCAM}} - (1-h) \odot \sum_{j,k}|\tilde{\mathcal{\bm{A}}}_{i,j,k}^{\text{HiResCAM}}|\right) \right) - (1 - y_i) \log \left( \bm{s}_i \right)
	\end{gather}
\end{theorem}
\begin{proof} For any logit $i$, binary cross-entropy loss is defined as:
	\begin{align}
		H(\theta, \mathcal{D}) &= -\left[ y_i \log (\sigma(\bm{s}_i)) + (1 - y_i) \log (1 - \sigma(\bm{s}_i)) \right] \\
		&= -\left[ y_i \log \left(\sigma\left(\sum_{j,k}\tilde{\mathcal{\bm{A}}}_{i,j,k}^{\text{HiResCAM}}\right) \right) + (1 - y_i) \log \left(1 - \sigma\left(\sum_{j,k}\tilde{\mathcal{\bm{A}}}_{i,j,k}^{\text{HiResCAM}}\right) \right) \right]
	\end{align}
	Similar to Theorem \ref{ace}, we can break down each HiResCAM to foreground and background components. For a non-target HiResCAM, anything that is not the target is inherently the background. Therefore, it's definition is retrained.
	\begin{gather}
		H(\theta, \mathcal{D}) = - y_i \log \left(\sigma\left(h \odot \sum_{j,k}\tilde{\mathcal{\bm{A}}}_{i,j,k}^{\text{HiResCAM}} - (1-h) \odot \sum_{j,k}\tilde{\mathcal{\bm{A}}}_{i,j,k}^{\text{HiResCAM}}\right) \right) - (1 - y_i) \log \left( \bm{s}_i \right)
	\end{gather}
	The current formulation motivates activating either the foreground or the background for positive classification, and motivates de-activating every pixel of the non-positive class. We penalize activation on the background for the positive class only:
	\begin{gather}
		H(\theta, \mathcal{D}) = - y_i \log \left(\sigma\left(h \odot \sum_{j,k}\tilde{\mathcal{\bm{A}}}_{i,j,k}^{\text{HiResCAM}} - (1-h) \odot \sum_{j,k}|\tilde{\mathcal{\bm{A}}}_{i,j,k}^{\text{HiResCAM}}|\right) \right) - (1 - y_i) \log \left( \bm{s}_i \right)
	\end{gather}
	This gives us the ablated binary cross-entropy formulation.
\end{proof}
